\LARGE \bf Preparation of Papers for IEEE Sponsored Conferences \& Symposia*

### 2017-02-08 implementation.tex
.

The implementation was prototyped as S3CAM-R, an extension to our
previously mentioned tool S3CAM (\chapref{case}). OLS regression
routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
Python module for machine learning. SAL~\cite{SAL-SRI}
with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
solver.

@@

We tabulate our preliminary evaluation in \tabref{res-rel}. We used
few of the benchmarks described in \chapref{case}, including the Van
der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
benchmark. As before, we ran S3CAM-R $10$ times with different seeds
and averaged the results. We tabulate both the total time taken and
the time taken by SAL to compute the counter-example and compare
against a newer implementation of S3CAM. Note that different
abstraction parameters are used for S3CAM and S3CAM-R, due to the
differences in which they operate.

\begin{table*}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in seconds (rounded off
to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
Benchmark & BMC & S3CAM-R & S3CAM\\
\midrule
Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
Brusselator               &$<$1  & $4$    & $2$\\
Lorenz                    &$<$1 & $122$  & $35$\\
%B.Ball                    &  &   & \\
Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
Bouncing Ball             &-  &$TO$   & $450$\\

\bottomrule
\end{tabular}
\end{table*}

>>>>

The implementation was prototyped as S3CAM-R, an extension to our
previously mentioned tool S3CAM \cite{zutshi2014multiple}. OLS regression
routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
Python module for machine learning. SAL~\cite{SAL-SRI}
with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
solver.

@@

We tabulate our preliminary evaluation in \tabref{res-rel}. We used
few of the benchmarks described in \cite{zutshi2014multiple}, including the Van
der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
benchmark. As before, we ran S3CAM-R $10$ times with different seeds
and averaged the results. We tabulate both the total time taken and
the time taken by SAL to compute the counter-example and compare
against a newer implementation of S3CAM. Note that different
abstraction parameters are used for S3CAM and S3CAM-R, due to the
differences in which they operate.

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in seconds (rounded off
to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
Benchmark & BMC & S3CAM-R & S3CAM\\
\midrule
Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
Brusselator               &$<$1  & $4$    & $2$\\
Lorenz                    &$<$1 & $122$  & $35$\\
%B.Ball                    &  &   & \\
Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
Bouncing Ball             &-  &$TO$   & $450$\\

\bottomrule
\end{tabular}
\end{table}

### 2017-02-08 prelims.tex
.

We define the PWA transition system as a transition system
(ref.~\secref{DiscSys}) $\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
where each transition $\tau \in \scrT$, is associated with a
transition relation of the form
\[
    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,f_{\tau}(\pVar)) \;|\; g_\tau(\pVar) \land g'_\tau(f_{\tau}(\pVar))}
\]
where $f_{\tau}$ is an affine map relating the states and $g_{\tau}$,
$g'_{\tau}$ are affine guards (pre and post conditions) on the states.
The assertion on the initial states $\Theta$ is also affine.

>>>>

We define the PWA transition system as a transition system
%(ref.~\secref{DiscSys})
$\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
where each transition $\tau \in \scrT$, is associated with a
transition relation of the form
\[
    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,f_{\tau}(\pVar)) \;|\; g_\tau(\pVar) \land g'_\tau(f_{\tau}(\pVar))}
\]
where $f_{\tau}$ is an affine map relating the states and $g_{\tau}$,
$g'_{\tau}$ are affine guards (pre and post conditions) on the states.
The assertion on the initial states $\Theta$ is also affine.

### 2017-02-08 pwarel.tex
.

We now briefly describe the approach. Given a numerical simulator
$\simulate$ and an initial abstraction defined by a quantization
function $\quant_\epsilon$ and a time step $\Delta$, we use
scatter-and-simulate (parameterized by number of samples $n$ as
described in \chapref{abs}) to explore the abstract graph
$\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
number of abstract states $C$ (or cells) and edges $(C,C')$ iff $C
\areach{\Delta} C'$.

Instead of using a CEGAR like loop (as in \chapref{abs}), we use the
generated trajectory segments to learn quantitative models describing
the local behavior of the system. These models are defined by a set of
relations $R \subseteq \HybridStates \times \HybridStates$ for each
edge of the reachability graph.

@@

Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
trajectory segment between the respective cells. The graph abstraction
$G$ only states that there exists a state $\x \in C$ from which the
system can evolve to a future state $\x' \in C'$. To increase the
precision we iteratively refined the abstraction by state
splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
which non-deterministically describe relations between $\x \in C$ and $\x'
\in C'$. This is illustrated in \figref{enriched-edge} (compare with
\figref{segtraj}).

@@

The existential abstraction relation in \chapref{abs} was defined as
follows:
\[
%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
\]
The abstract relation $\areach{t}$ can be enriched by incorporating
the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
system, such a relation can be rarely represented using an exact affine map.
This is due to the presence of non-linear and hybrid behaviors. But,
an affine map can always be estimated with an error.

@@

\end{scope}
\begin{scope}[xshift=0,yshift=3.0cm,scale=0.5]
    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
    \implies A_1\vx + \vb_1 + \delta_1$};
\end{scope}

@@

\end{scope}
\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
    \node at (2,0) {\begin{minipage}{\linewidth}
    \footnotesize \begin{equation*}
        \scr{T} = \left\{
        \begin{array}{ll}
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
        \end{array}
    \right.
\end{equation*}
\end{minipage}};
\end{scope}

@@

\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
    \node at (4,0) {\begin{minipage}{\linewidth}
        \footnotesize \begin{equation*}
    \scr{T} = \left\{
        \begin{array}{ll}
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
        \end{array}
    \right.
\end{equation*}
\end{minipage}};
\end{scope}

@@

Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
\textit{similar} if their trajectory segments have the same sequence
of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
\[
    \forall i\in\setof{1 \ldots k}.
    start(\pi_{t_i}) \in C_i
    \land start(\pi'_{t_i}) \in C_i
    \land end(\pi_{t_i}) \in C_{i+1}
    \land end(\pi'_{t_i}) \in C_{i+1}
\]

>>>>

We now briefly describe the approach. Given a numerical simulator
$\simulate$ and an initial abstraction defined by a quantization
function $\quant_\epsilon$ and a time step $\Delta$, we use
scatter-and-simulate (parameterized by number of samples $n$ as
described in \cite{zutshi2014multiple}) to explore the abstract graph
$\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
number of abstract states $C$ (or cells) and edges $(C,C')$ iff $C
\areach{\Delta} C'$.

Instead of using a CEGAR like loop (as in \cite{zutshi2014multiple}), we use the
generated trajectory segments to learn quantitative models describing
the local behavior of the system. These models are defined by a set of
relations $R \subseteq \HybridStates \times \HybridStates$ for each
edge of the reachability graph.

@@

Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
trajectory segment between the respective cells. The graph abstraction
$G$ only states that there exists a state $\x \in C$ from which the
system can evolve to a future state $\x' \in C'$. To increase the
precision we iteratively refined the abstraction by state
splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
which non-deterministically describe relations between $\x \in C$ and $\x'
\in C'$. This is illustrated in \figref{enriched-edge}.
%(compare with \figref{segtraj}).

@@

The existential abstraction relation in \cite{zutshi2014multiple} was defined as
follows:
\[
%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
\]
The abstract relation $\areach{t}$ can be enriched by incorporating
the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
system, such a relation can be rarely represented using an exact affine map.
This is due to the presence of non-linear and hybrid behaviors. But,
an affine map can always be estimated with an error.

@@

\end{scope}
\begin{scope}[xshift=0,yshift=2.0cm,scale=0.5]
    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
    \implies A_1\vx + \vb_1 + \delta_1$};
\end{scope}

@@

\end{scope}
\begin{scope}[xshift=0,yshift=2.5cm,scale=0.5]
    \node at (2,0) {\begin{minipage}{\linewidth}
    \footnotesize \begin{equation*}
        \scr{T} = \left\{
        \begin{array}{ll}
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
        \end{array}
    \right.
\end{equation*}
\end{minipage}};
\end{scope}

@@

\begin{scope}[xshift=0,yshift=2.5cm,scale=0.5]
    \node at (4,0) {\begin{minipage}{\linewidth}
        \footnotesize \begin{equation*}
    \scr{T} = \left\{
        \begin{array}{ll}
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
        \end{array}
    \right.
\end{equation*}
\end{minipage}};
\end{scope}

@@

Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
\textit{similar} if their trajectory segments have the same sequence
of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
\begin{align*}
    \forall i\in\setof{1 \ldots k}.
    & start(\pi_{t_i}) \in C_i          
    \land start(\pi'_{t_i}) \in C_i \land  \\
    & end(\pi_{t_i}) \in C_{i+1}  
    \land end(\pi'_{t_i}) \in C_{i+1}
\end{align*}
